---
title: "Code for prediction error"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Code for prediction error}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(JMtwostage)
library(survival)
```
This vignette demonstrates how to use the `predError()` function to compute prediction error measures such as the Brier score for survival models fitted under various missing data approaches.

The example illustrates the process of:
* Fitting a joint model using the `jmwMI()` function, which combines multiple imputation and      inverse probability weighting for longitudinal and survival data.
* Specifying a time-dependent Cox model with longitudinal biomarkers included as predictors.
* Calculating prediction error using the `predError()` function, which internally uses the `brier()` function to evaluate model performance at multiple time points.

**Note**: Ensure that the required input datasets (`long_data`, `surv_data`) are properly preprocessed and available in your environment before running the example code.
```{r eval=FALSE}
predError <- function(fitted.object,
                      model_cov,
                      newdata = NULL,
                      nimp,
                      times = seq(2, 10, 0.5),...) {

  if (fitted.object$method == "MI" || fitted.object$method == "wMI") {
    fitted_model <- fitted.object$fitted_model[[nimp]]
  } else if (fitted.object$method %in% c("CC", "2s", "LOCF")) {
    fitted_model <- fitted.object$fitted_model[[1]]
  }

  if (fitted.object$method %in% c("MI", "wMI")) {
    data1 <- fitted.object$complete_data[[nimp]]
  } else {
    data1 <- fitted.object$complete_data
  }
  complete_data <- data1

  if (fitted.object$method == "wMI") {
    cfit <- coxph(
      as.formula(paste0("Surv(tstart, tstop, new_event2) ~ ", paste0(model_cov, collapse = '+'))),
      weights = complete_data$product_W_ij,
      id = complete_data[["ID"]],
      data = complete_data
    )
  } else {
    cfit <- coxph(
      as.formula(paste0("Surv(tstart, tstop, new_event2) ~ ", paste0(model_cov, collapse = '+'))),
      id = complete_data[["ID"]],
      data = complete_data
    )
  }

  if (is.null(newdata)) {
    newdata <- complete_data
  }

  brier_results <- brier(cfit, newdata = newdata, times = times,...)
  return(brier_results)
}

model_wMI <- jmwMI(
  ldata = long_data,
  sdata = surv_data,
  timeDep = c("marker_1", "marker_2", "marker_3"),
  impModel = list(
    marker_1 ~ Z_1 + Z_2 + Time + (1 | ID),
    marker_2 ~ Z_1 + Z_2 + Time + (1 | ID),
    marker_3 ~ Z_1 + Z_2 + Time + (1 | ID)
  ),
  ipwModel = list(
    ~ Z_1 + Time + (1 | ID),
    ~ Z_1 + Time + (1 | ID),
    ~ Z_1 + Time + (1 | ID)
  ),
  visitTime = "Time",
  coxModel = Surv(survival_time, survival_status) ~ Z_1 + Z_2 +
    td(marker_1) + td(marker_2) + td(marker_3),
  model = "Cox",
  id = "ID"
)

predError(
  fitted.object = model_wMI,
  model_cov = c("Z_1", "Z_2", "marker_1", "marker_2", "marker_3"),
  nimp = 1,
  times = seq(2, 10, 0.5)
)
```
